### 2010-04-15 doc/paper.tex
updated the docs a bit and added a .gitignore

\section*{Abstract}
\small{This paper outlines two implementations of the VPM-B dive algorithm: a reference implementation for
  generating dive tables, and a proof of concept real-time decompression algorithm, suited for
  use in a dive computer.

@@

 Today, there are many algorithms used to help divers safely avoid decompression sickness. They can either be
 used to calculate dive tables, which allow the diver to match their dive profile to a decompression schedule,
 or be used in a dive computer, which track the time and depth of the dive in real time, and use that information
 to calculate the decompression schedule. While there are already programs that can be used to generate the
 dive tables\footnote{VPMOpen {\color{blue}\uline{\href{http://webspace.webring.com/people/fv/vpmopen/}{http://webspace.webring.com/people/fv/vpmopen/}}}\\ and the original VPM-B program by Erik Baker \\{\color{blue}\uline{\href{ftp://decompression.org/Baker/VPM-B\%20Fortran\%20Source\%20Code.txt}{ftp://decompression.org/Baker/VPM-B\%20Fortran\%20Source\%20Code.txt}}} are two such examples}, the only programs that do real time analysis for decompression are proprietary,
embedded systems for dive computers. The closed nature of these systems means that they may not be following
the algorithm correctly, or other bugs in the system may make them less reliable than they could be, putting
the safety of divers at risk.

@@

This section provides a basic introduction to scuba diving, and what you need to know about decompression, to
help readers understand what will be discussed in this paper. 
When you scuba dive, you breath in compressed gasses \footnote{Often air, but other gas mixes are used
for more complicated dives}. These gas mixes contain oxygen, and almost always other gasses \footnote{Pure oxygen
is usually not used for safety reasons(Oxygen Toxicity)}. Nitrogen and Helium are the most commonly used
other gasses, and when you breathe them under pressure, they start to form bubbles in body tissues. If the diver rises to quickly to the surface, these bubbles will expand, and can cause many serious medical problems, including death. This can
be avoided if the diver rises to specific depths, and waits a certain amount of time for the gasses to leave the
tissue (offgassing). Due to the risk, many models have been created that allow divers to calculate the depth and
time required to safely decompress before returning to the surface. The VPM-B algorithm is one such model that
has been chosen for this project, based on the large amount of documentation, and availability of example code.

@@

\subsubsection{VPM}
 The VPM algorithm starts by adopting a model of body tissues developed by Albert A. B\"uhlmann. This model
uses 16 tissue compartments to model the effect the dissolved gasses have on human tissue. During the
normal part of the dive, gas is added to the compartments\cite{Buhlmann84} (on-gassing\glossary{name={On-Gassing}, description={The process of adding inert gasses to body tissues}}), and during the final ascent,
they're released from the tissue (off-gassing\glossary{name={Off-Gassing}, description={The process of releasing inert gasses from body tissues}}). Many different compartments are needed because different
tissues absorb and release gasses at different rates (as evidenced by the Helium\_Half\_Time and
Nitrogen\_Half\_Time arrays in the program).

@@

 Once the Python version was up and running, it was time to make many changes to the code to make it more
in line with the programming practices taught at Carleton University. The original program was written for FORTRAN 77,
and followed many of the practices common to FORTRAN\glossary{name={FORTRAN}, description={A programming language, popular with scientific researchers, but
not often taught to programmers these days}} programs (GOTOs, global state used for all long term storage,
custom file formats for input data), which would be very hard for modern day programmers to follow. The program
was made object oriented\footnote{There are definite instances of the ``God Object'' anti-pattern, but at least the state is encapsulated, and in a position to be refactored}, by ensuring that all the global state was encapsulated in a DiveState object, and all functions that would modify this state belonged to the object.
The GOTOs were turned into normal loops with break statements, extremely long functions were broken into smaller
pieces, and repeating code was extracted to stand alone methods. Finally,  the input and output file formats were changed
to use JSON\glossary{name={JSON}, description={JavaScript Object Notation, a lightweight data representation language
designed to be easier to write than XML}}
, as it is a standard format like XML\glossary{name={XML}, description={Extensible Markup Language, a language used
to structure data for easy computer reading and writing}}, that is supported by just about every programming language \footnote{though not for FORTRAN, so a translator was written to let the original program load the new input files}.

@@

Finally, for this proof of concept algorithm, there are a couple of restrictions for the
decompression: first, the diver will have to signal that they want to decompress \footnote{by pushing a button
on a real computer we'll assume}, and second, that they obey the decompression 
schedule without skipping stops, or spending too
little time at them. While more advanced dive computers are able to give a best effort estimation of how to
continue the decompression schedule after a missed stop, there wasn't enough information about
how to safely handle missed stops, so it has been left out rather than estimated.

@@

Finally, just getting a lot of programmers to read the code, point out problems, and criticize the program
 would be a great step in improving the quality of the code. Since C programs can suffer from
vary strange and random bugs having other people help improve the program would go a long way towards making this
as bug free, and feature complete as possible.

@@

\subsection{Input}
\subsubsection{Main Part}
%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                 &  Type                         &  Description                                                          \\
\hline
 ``desc''             &  string                       &  description of the dive which will appear at the top of the output.  \\
\hline
 ``num\_gas\_mixes''  &  integer                      &  the number of gas mixes held in the ``gasmix\_summary'' array.       \\
\hline
 ``gasmix\_summary''  &  array[hash\{string:float\}]  &  gas mixes carried by the diver, in the order they will be used.      \\
\hline
\end{longtable}
%\end{center}
%\end{table}

@@

\subsubsection{Profiles}
%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                &  Type         &  Description                                          \\
\hline
 ``profile\_codes''  &  array[hash]  &  descent, constant depth, and ascent to be performed  \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                 &  Type     &  Description                                         \\
\hline
 ``profile\_code''    &  integer  &  must be 1 for descent or ascent                               \\
\hline
 ``starting\_depth''  &  float    &  depth in FSW or MSW where the dive will be started  \\
\hline
 ``ending\_depth''    &  float    &  depth in FSW or MSW where the dive will stop        \\
\hline
 ``rate''             &  float    &  FSW/min or MSW/min the diver will descend at        \\
\hline
 ``gasmix''           &  integer  &  gasmix being used                                   \\
\hline
 ``setpoint''         &  float    &  no clue                                             \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                 &  Type     &  Description                                                \\
\hline
 ``profile\_code''                    &  integer  &  must be 2 for constant depth                               \\
\hline
 ``depth''                            &  float    &  depth where the dive will take place                       \\
\hline
 ``run\_time\_at\_end\_of\_segment''  &  float    &  what time (in minutes) the dive at the ``depth'' will end  \\
\hline
 ``gasmix''                           &  integer  &  gasmix being used                                          \\
\hline
 ``setpoint''                         &  float    &  no clue                                                    \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                        &  Type         &  Description                                                                          \\
\hline
 ``profile\_code''                           &  integer      &  must be 99 for ascent                                                                \\
\hline
 ``number\_of\_ascent\_parameter\_changes''  &  integer      &  number of changes during the ascent (a change would be the step size, rate, gasmix)  \\
\hline
 ``ascent\_summary''                         &  array[hash]  &  ascent profiles to be followed                                                       \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                 &  Type            &  Description                                                         \\
\hline
 ``starting\_depth''  &  float           &  depth where the ascent profile starts                               \\
\hline
 ``gasmix''           &  integer         &  gasmix summary being used                                           \\
\hline
 ``rate''             &  negative float  &  the rate of ascent in FSW/min or MSW/min (must be negative number)  \\
\hline
 ``step\_size''       &  float           &  distance between each decompression stop                            \\
\hline
 ``setpoint''         &  float           &  no clue                                                             \\
\hline
\end{longtable}
%\end{center}

@@

\hline
 Name                                   &  Type    &  Description \\                                                         \hline
 ``repetitive\_code''                  &  integer  &  do another dive (1), or stop (0)             \\
\hline
 ``surface\_interval\_time\_minutes''  &  float    &  time spent at surface before a repeat dive.  \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                   &  Type    &  Description                                                                                                  \\
\hline
 ``Altitude\_of\_Dive''                 &  float   &  height in meters or feet the dive will occur at                                                              \\
\hline
 ``Diver\_Acclimatized\_at\_Altitude''  &  string  &  ``yes'', or ``no'', based on whether the diver has spent enough time at the altitude to become acclimatized  \\
\hline
 ``Starting\_Acclimatized\_Altitude''   &  float   &  altitude diver is acclimatized at                                                                            \\
\hline
 ``Ascent\_to\_Altitude\_Hours''        &  float   &  time in hours the diver took to get to the diving altitude                                                   \\
\hline
 ``Hours\_at\_Altitude\_Before\_Dive''  &  float   &  time the diver has spent at the altitude of the dive                                                         \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                  &  Type                                                                                             &  Description                                                                              \\
\hline
 ``Units''                             &  string                                                                                           &  either ``fsw'' or ``msw'' if the dive is in feet of salt water, or meters of salt water  \\
\hline
 ``Altitude\_Dive\_Algorithm''         &  string                                                                                           &  ``ON'' or ``OFF'' based on whether we're doing an altitude dive.                         \\
\hline
 ``Minimum\_Deco\_Stop\_Time''         &  float                                                                                            &  minimum time the diver will spend at a decompression stop                                \\
\hline
 ``Critical\_Radius\_N2\_Microns''     &  float                                                             & Critical nitrogen bubble size                                                                                         \\
\hline
 ``Critical\_Radius\_He\_Microns''     &  float                                                               &  Critical helium bubble size                                                                                         \\
\hline
 ``Critical\_Volume\_Algorithm''       &  string  &  ``ON'' or ``OFF'' based on whether or not we want to run the critical volume algorithm                                                                                         \\
\hline
 ``Crit\_Volume\_Parameter\_Lambda''   &  float                                                                                            &  Determines the safe bubble growth limit (you probably don't want to touch it)                                                                                     \\
\hline
 ``Gradient\_Onset\_of\_Imperm\_Atm''  &  float                                                                                            &  Must be between 5 and 10, but you should probably leave it at the default 8.2                                                                                     \\
\hline
 ``Surface\_Tension\_Gamma''           &  float                                                                                            &  surface tension of the bubble                                                            \\
\hline
 ``Skin\_Compression\_GammaC''         &  float                                                                                            &  I think this is the ``skin'' of the bubble                                               \\
\hline
 ``Regeneration\_Time\_Constant''      &  float                                                                                            &  I think this is the time a bubble takes to return to its original size                   \\
\hline
 ``Pressure\_Other\_Gases\_mmHg''      &  float                                                                                            &  other gases (carbon dioxide, water vapor and oxygen).                                    \\
\hline
\end{longtable}
%\end{center}

@@

\subsection{Dive State}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name  &  Type  &  Description \\
\hline
Starting\_Depth & double & Starting depth for the current dive profile \\
\hline
Ending\_Depth & double &  Ending depth of the profile\\
\hline
Rate & double & Rate of depth change for ascents and descents. Either calculated in real time, or set ahead of time \\
\hline
Depth & double &  Current depth of the diver\\
\hline
Last\_Depth & double & Depth of the last iteration of the real dive time loop. \\
\hline
Profile\_End\_Time & double & The time the constant depth profile will end. Called Run\_Time\_End\_of\_Segment in implementations.\\
\hline
Run\_Time & double & Current run time of the dive \\
\hline
Depth\_Start\_of\_Deco\_Zone & double & Depth that the diver can decompress at \\
\hline
Step\_Size & double & After the first decompression stop, this determines the depths that the decompression stops will take place at. \\
\hline
Deco\_Stop\_Depth & double & During decompression, this is used to set the next stop the diver must ascend to. \\
\hline
First\_Stop\_Depth & double &  The depth the first decompression stop took place at.\\
\hline
Schedule\_Converged & boolean & Check to see if the Phase\_Volume\_Time is less than or equal to a minute difference from the Last\_Phase\_Volume\_Time. If true, we can stop running the Critical\_Volume\_Algorithm \\
\hline
Critical\_Volume\_Algorithm\_Off & boolean & Flag to tell us not to run the Critical\_Volume\_Algorithm. \\
\hline
Deco\_Phase\_Volume\_Time & double & Used when applying the Critical\_Volume\_Algorithm \\
\hline
Helium\_Pressure & double[16] & The B\"uhlmann compartments for helium. \\
\hline
Nitrogen\_Pressure & double[16] & The B\"uhlmann compartments for nitrogen. \\
\hline
*decomp\_stops & decompression\_stops & Array for storing the decompression stop depths, times, and directions, 
calculated when the diver decides to decompress. \\
\hline
decomp\_stop\_index & int & Points to the current decompression stop in the decomp\_stops array. \\
\hline

>>>>

\section*{Abstract}
\small{This paper outlines two implementations of the VPM-B dive algorithm: a reference implementation for
  generating dive tables, and a proof of concept, real-time decompression algorithm, suited for
  use in a dive computer.

@@

 Today, there are many algorithms used to help divers safely avoid decompression sickness. They can either be
 used to calculate dive tables, which allow the diver to match their dive profile to a decompression schedule,
 or be used in a dive computer, which tracks the time and depth of the dive in real time, and uses that information
 to calculate the decompression schedule. While there are already programs that can be used to generate the
 dive tables\footnote{VPMOpen {\color{blue}\uline{\href{http://webspace.webring.com/people/fv/vpmopen/}{http://webspace.webring.com/people/fv/vpmopen/}}}\\ and the original VPM-B program by Erik Baker \\{\color{blue}\uline{\href{ftp://decompression.org/Baker/VPM-B\%20Fortran\%20Source\%20Code.txt}{ftp://decompression.org/Baker/VPM-B\%20Fortran\%20Source\%20Code.txt}}} are two such examples}, the only programs that do real time analysis for decompression are proprietary,
embedded systems for dive computers. The closed nature of these systems means that they may not be following
the algorithm correctly, or other bugs in the system may make them less reliable than they could be, putting
the safety of divers at risk.

@@

This section provides a basic introduction to scuba diving, and what you need to know about decompression to
 understand what will be discussed in this paper. 
When you scuba dive, you breath in compressed gasses \footnote{Often air, but other gas mixes are used
for more complicated dives}. These gas mixes contain oxygen, and almost always other gasses \footnote{Pure oxygen
is usually not used for safety reasons(Oxygen Toxicity)}. Nitrogen and Helium are the most commonly used
other gasses, and when you breathe them under pressure, they start to form bubbles in body tissues. If the diver rises to quickly to the surface, these bubbles will expand, and can cause many serious medical problems, including death. This can
be avoided if the diver rises to specific depths, and waits a certain amount of time for the gasses to leave the
tissue (offgassing). Due to the risk, many models have been created that allow divers to calculate the depth and
time required to safely decompress before returning to the surface. The VPM-B algorithm is one such model that
has been chosen for this project, based on the large amount of documentation, and availability of example code.

@@

\subsubsection{VPM}
 The VPM algorithm starts by adopting a model of body tissues developed by Albert A. B\"uhlmann. This model
uses 16 tissue compartments to model the effect the dissolved gasses have on human tissues. During the
normal part of the dive, gas is added to the compartments\cite{Buhlmann84} (on-gassing\glossary{name={On-Gassing}, description={The process of adding inert gasses to body tissues}}), and during the final ascent,
they're released from the tissue (off-gassing\glossary{name={Off-Gassing}, description={The process of releasing inert gasses from body tissues}}). Many different compartments are needed because different
tissues absorb and release gasses at different rates (as evidenced by the Helium\_Half\_Time and
Nitrogen\_Half\_Time arrays in the program).

@@

 Once the Python version was up and running, it was time to make many changes to the code to make it more
in line with the programming practices taught at Carleton University. The original program was written for FORTRAN 77,
and followed many of the practices common to FORTRAN\glossary{name={FORTRAN}, description={A programming language, popular with scientific researchers, but
not often taught to programmers these days}} programs (GOTOs, global state used for all long term storage,
custom file formats for input data), which would be very hard for modern day programmers to follow. The program
was made object oriented\footnote{There are definite instances of the ``God Object'' anti-pattern, but at least the state is encapsulated, and in a position to be refactored}, by ensuring that all the global state was encapsulated in a DiveState object, and all functions that would modify this state belonged to the object.
The GOTOs were turned into normal loops with break statements, extremely long functions were broken into smaller
pieces, and repeating code was extracted to stand alone methods. Finally,  the input and output file formats were changed
to use JSON,\glossary{name={JSON}, description={JavaScript Object Notation, a lightweight data representation language
designed to be easier to write than XML}}
 as it is a standard format like XML\glossary{name={XML}, description={Extensible Markup Language, a language used
to structure data for easy computer reading and writing}}, that is supported by just about every programming language \footnote{Though not FORTRAN, so a translator was written to let the original program load the new input files}.

@@

Finally, for this proof of concept algorithm, there are a couple of restrictions for the
decompression: first, the diver will have to signal that they want to decompress \footnote{By pushing a button
on a real computer we'll assume}, and second, that they obey the decompression 
schedule without skipping stops, or spending too
little time at them. While more advanced dive computers are able to give a best effort estimation of how to
continue the decompression schedule after a missed stop, there wasn't enough information about
how to safely handle missed stops, so it has been left out rather than estimated.

@@

Finally, just getting a lot of programmers to read the code, point out problems, and criticize the program
 would be a great step in improving the quality of the code. Since C programs can suffer from
very strange and random bugs having other people help improve the program would go a long way towards making this
as bug free, and feature complete as possible.

@@

\subsection{Input}
\subsubsection{Main Part}
%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                 &  Type                         &  Description                                                          \\
\hline
 ``desc''             &  string                       &  Description of the dive which will appear at the top of the output.  \\
\hline
 ``num\_gas\_mixes''  &  integer                      &  The number of gas mixes held in the ``gasmix\_summary'' array.       \\
\hline
 ``gasmix\_summary''  &  array[hash\{string:float\}]  &  Gas mixes carried by the diver, in the order they will be used.      \\
\hline
\end{longtable}
%\end{center}
%\end{table}

@@

\subsubsection{Profiles}
%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                &  Type         &  Description                                          \\
\hline
 ``profile\_codes''  &  array[hash]  &  Descents, ascents, and constant depth segments to be performed  \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                 &  Type     &  Description                                         \\
\hline
 ``profile\_code''    &  integer  &  Must be 1 for descent or ascent                               \\
\hline
 ``starting\_depth''  &  float    &  Depth in FSW or MSW where the dive will be started  \\
\hline
 ``ending\_depth''    &  float    &  Depth in FSW or MSW where the dive will stop        \\
\hline
 ``rate''             &  float    &  FSW/min or MSW/min the diver will descend at        \\
\hline
 ``gasmix''           &  integer  &  Gasmix being used                                   \\
\hline
 ``setpoint''         &  float    &  Not Used (mentioned in Baker's code, so it's reproduced here)                                             \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                 &  Type     &  Description                                                \\
\hline
 ``profile\_code''                    &  integer  &  Must be 2 for constant depth                               \\
\hline
 ``depth''                            &  float    &  Depth where the dive will take place                       \\
\hline
 ``run\_time\_at\_end\_of\_segment''  &  float    &  What time (in minutes) the dive at the ``depth'' will end  \\
\hline
 ``gasmix''                           &  integer  &  Gasmix being used                                          \\
\hline
 ``setpoint''                         &  float    &  Not Used                                                    \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                        &  Type         &  Description                                                                          \\
\hline
 ``profile\_code''                           &  integer      &  Must be 99 for ascent                                                                \\
\hline
 ``number\_of\_ascent\_parameter\_changes''  &  integer      &  Number of changes during the ascent (a change would be the step size, rate, gasmix)  \\
\hline
 ``ascent\_summary''                         &  array[hash]  &  Ascent profiles to be followed                                                       \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                 &  Type            &  Description                                                         \\
\hline
 ``starting\_depth''  &  float           &  Depth where the ascent profile starts                               \\
\hline
 ``gasmix''           &  integer         &  Gasmix summary being used                                           \\
\hline
 ``rate''             &  negative float  &  The rate of ascent in FSW/min or MSW/min (must be negative number)  \\
\hline
 ``step\_size''       &  float           &  Distance between each decompression stop                            \\
\hline
 ``setpoint''         &  float           &  Not Used                                                             \\
\hline
\end{longtable}
%\end{center}

@@

\hline
 Name                                   &  Type    &  Description \\                                                         \hline
 ``repetitive\_code''                  &  integer  &  Do another dive (1), or stop (0)             \\
\hline
 ``surface\_interval\_time\_minutes''  &  float    &  Time spent at surface before a repeat dive.  \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                   &  Type    &  Description                                                                                                  \\
\hline
 ``Altitude\_of\_Dive''                 &  float   &  Height in meters or feet the dive will occur at                                                              \\
\hline
 ``Diver\_Acclimatized\_at\_Altitude''  &  string  &  ``yes'', or ``no'', based on whether the diver has spent enough time at the altitude to become acclimatized  \\
\hline
 ``Starting\_Acclimatized\_Altitude''   &  float   &  Altitude diver is acclimatized at                                                                            \\
\hline
 ``Ascent\_to\_Altitude\_Hours''        &  float   &  Time in hours the diver took to get to the diving altitude                                                   \\
\hline
 ``Hours\_at\_Altitude\_Before\_Dive''  &  float   &  Time the diver has spent at the altitude of the dive                                                         \\
\hline
\end{longtable}
%\end{center}

@@

%\begin{center}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name                                  &  Type                                                                                             &  Description                                                                              \\
\hline
 ``Units''                             &  string                                                                                           &  Either ``fsw'' or ``msw'' if the dive is in feet of salt water, or meters of salt water  \\
\hline
 ``Altitude\_Dive\_Algorithm''         &  string                                                                                           &  ``ON'' or ``OFF'' based on whether we're doing an altitude dive.                         \\
\hline
 ``Minimum\_Deco\_Stop\_Time''         &  float                                                                                            &  Minimum time the diver will spend at a decompression stop                                \\
\hline
 ``Critical\_Radius\_N2\_Microns''     &  float                                                             & Critical nitrogen bubble size                                                                                         \\
\hline
 ``Critical\_Radius\_He\_Microns''     &  float                                                               &  Critical helium bubble size                                                                                         \\
\hline
 ``Critical\_Volume\_Algorithm''       &  string  &  ``ON'' or ``OFF'' based on whether or not we want to run the critical volume algorithm                                                                                         \\
\hline
 ``Crit\_Volume\_Parameter\_Lambda''   &  float                                                                                            &  Determines the safe bubble growth limit (you probably don't want to touch it)                                                                                     \\
\hline
 ``Gradient\_Onset\_of\_Imperm\_Atm''  &  float                                                                                            &  Must be between 5 and 10, but you should probably leave it at the default 8.2                                                                                     \\
\hline
 ``Surface\_Tension\_Gamma''           &  float                                                                                            &  Surface tension of the bubble                                                            \\
\hline
 ``Skin\_Compression\_GammaC''         &  float                                                                                            &  Constant representing the permeable ``skin'' of the bubble                                               \\
\hline
 ``Regeneration\_Time\_Constant''      &  float                                                                                            &  The time a bubble takes to return to its original size                   \\
\hline
 ``Pressure\_Other\_Gases\_mmHg''      &  float                                                                                            &  Other gases (carbon dioxide, water vapor and oxygen).                                    \\
\hline
\end{longtable}
%\end{center}

@@

\subsection{Dive State}
\begin{longtable}{|l|l|p{5cm}|}
\hline
 Name  &  Type  &  Description \\
\hline
Starting\_Depth & double & Starting depth for the current dive profile \\
\hline
Ending\_Depth & double &  Ending depth of the profile\\
\hline
Rate & double & Rate of depth change for ascents and descents. Either calculated in real time, or set ahead of time \\
\hline
Depth & double &  Current depth of the diver\\
\hline
Last\_Depth & double & Depth of the last iteration of the real dive time loop. \\
\hline
Profile\_End\_Time & double & The time the constant depth profile will end. Called Run\_Time\_End\_of\_Segment in implementations.\\
\hline
Run\_Time & double & Current run time of the dive \\
\hline
Depth\_Start\_of\_Deco\_Zone & double & Depth that the diver can decompress at \\
\hline
Step\_Size & double & After the first decompression stop, this determines the depths that the decompression stops will take place at. \\
\hline
Deco\_Stop\_Depth & double & During decompression, this is used to set the next stop the diver must ascend to. \\
\hline
First\_Stop\_Depth & double &  The depth the first decompression stop took place at.\\
\hline
Schedule\_Converged & boolean & Check to see if the Phase\_Volume\_Time is less than or equal to a minute difference from the Last\_Phase\_Volume\_Time. If true, we can stop running the Critical\_Volume\_Algorithm \\
\hline
Critical\_Volume\_Algorithm\_Off & boolean & Flag to tell us not to run the Critical\_Volume\_Algorithm. \\
\hline
Deco\_Phase\_Volume\_Time & double & Used when applying the Critical\_Volume\_Algorithm \\
\hline
Helium\_Pressure & double[16] & The B\"uhlmann compartments for helium. \\
\hline
Nitrogen\_Pressure & double[16] & The B\"uhlmann compartments for nitrogen. \\
\hline
decomp\_stops & decompression\_stops[] & Array for storing the decompression stop depths, times, and directions, 
calculated when the diver decides to decompress. \\
\hline
decomp\_stop\_index & int & Points to the current decompression stop in the decomp\_stops array. \\
\hline
